% !TeX root = ../main.tex
\chapter*{Prologue} \chaptermark{Prologue}


In the depth of the night I have always felt most comfortable. The stillness and
quietness, but for the sounds of the wind and every so often a creak as a
building adjusts itself, lets me feel like I've some control over my thoughts.
The lack of stimulation allows my mind to breathe, and I can think clearly and
focus on the task at hand. That is at least until \textit{he} ruins it by
opening his mouth. 

\begin{dialogue}
    \item{"Do you think they're out?", Bill wondered aloud. As always he wasted
    words on things so obvious.
    To an external observer this might have seemed like a perfectly innocent and valid
    question. But, Bill and I have been at this a while and all the signs are
    there in front of him, same as me. To ask such a question fills me with
    worries about his competence for this job.}
    \item {"No, the bedroom window is open."}
    \item {"How'd you know that's the bedroom?"}
\end{dialogue}

I knew this would continue until daybreak if I let it. Bill has lived with me
long enough to know that I heard his question and was purposefully ignoring it.
I swear once this is done I'm leaving this city, and not just because of Bill...

Entering into a house whilst the inhabitants are home has its upsides and
downsides. I knew, for instance, that the item we've been asked to procure is
never far from its \textit{current} owner. The downside of course being that
looming danger of being heard, seen, attacked, or any number of horrible things
that someone will do to protect themselves when feeling threatened. I mulled
over the various details of the building as I moved towards the narrow alley off
to the side. I'm certain I would have been thinking more clearly if I wasn't
being stalked by my shadow of a partner.

Normally, we wouldn't take such a specific job as this, but when my fence,
Edith, specifically recommended me I was of course intrigued, I assumed that she
had some inkling to my interest in planeweaving. Otherwise, it would be very
bizarre to ask for Bill and me when other, much more experienced and reliable
thieves could have done. Sealing my
interest and quelling my suspicions about the nature of being handed a job
rather than apply (as was the norm), was the small fortune in gold which was the
prize offered by the person(s) making the request. The amount of gold would be enough for a new
start, and a clean, guiltless break. Bill would be giddy enough with his split
that he might not even notice my departure for a few days. I made a point of not
planning past this, whilst I personally had issues with Bill, he didn't make it
this far in life on dumb luck. If I planned, if I left a clue to my next move,
then I'm sure he could have followed if he was so inclined. He was an unknown
and I did not want to find out what he was capable of. Thus, my next move will
be a surprise, even to me. 

But first, Ellen, focus. 

\begin{dialogue}
    \item{"I'm going to go in through the top window. You wait here and watch
    the..." started Bill as he looked for a foothold to begin climbing.}
    \item {"You can't see it from here, but the windows all have metal bars
    across them" I interrupted.
    In all fairness, I have always been our scout for heists. Walking past
    several times in the past week allowed me to notice some things which
    darkness hid.}
    \item{"Oh, right. I'll go in through the front door then." snorted Bill sarcastically.}
    \item{"Look", I pointed to the long, thin, rectangular windows just out of
    reach.
     "They're used for ventilation, and it'll be a squeeze, but I think we both could shimmy through."}
    \item{"Good spot", Bill said as he jumped for the small ledge.}
\end{dialogue}

Swinging his legs up and holding himself alongside the window, he showed why he
was in this line of work. As if having practised this exact manoeuvre a thousand
times he reached into one of his many pockets and drew back his hands now coated
in a tacky substance. Placing his hands on the glass of the window he pulled.
Slowly, steadily, and taking care not to lose his balance on the thin ledge he
pulled the ventilation window open. I couldn't help but smile at this display
of ingenuity. 

Taking a step backwards I took a steadying breath, lunging forward, left foot on
the cobbled stone, I pushed upwards, right foot connecting with the wall and
continuing the momentum I reached for the ledge of the window. Without stopping,
I pulled myself upwards and, just like Bill had done a moment before, I swung my
legs and body upwards and sideways. However, with the window open and Bill
already through and into the house, I passed through and into the darkness of
the ground floor, silently.


Whilst my eyes were already used to the darkness, it was the smells that
took me by surprise, and required a moment to adjust.  

\begin{dialogue}
    \item{"Gods, that explains the ventilation." I whispered.}
    \item{The colour drained from Bill's face as he looked at me, "Did you know?"}
    \item{"I didn't know exactly, but I had a feeling."}
    \item{"You didn't think to fucking mention we were breaking into a fucking alchemist's house?", I could see the fear and anger in Bill, he couldn't decide how to feel right now.}
\end{dialogue}

I gave him a look to indicate this discussion wouldn't be had now. Bill,
normally a bit more curious before accepting a job was so blinded by the
potential payout he'd not questioned it. He has changed, I thought to myself. I
remember being absolutely enamoured by his mind, the thoughts he shared with me,
that made me see the world differently. For a moment, forgetting where I was, I
felt a deep sense of loss. 

It was Bill's hurried movements that brought me back to our current reality.
His instincts kicked in, he was checking for sentries, alarms and traps. I stood
still, this was his world, not mine. I understood the basics and the sciences of
planeweaving, but when it came to using these things for defence or to harm...
well, my brief and interrupted education hadn't taken me that far. Bill knew
these things as a survivor, and cared little for how they worked, and more how
to avoid being killed by them.

Whilst he was checking the room, I finally took a moment to look around and
contemplate the sights and the smells of this room. More of a laboratory really,
now that I could see it. The room was long and tunnel like, filled with glass vials, beakers,
cauldrons suspended over smouldering coals and a heavy looking wooden door at
the far end. Novice as I am, I could still
discern the sweet smells of herbs used in healing potions, the harsh grassy
smell of poisons, and the distinct static in the air that comes with
planeweaving. There was magic in this room that was being bound against its
will, reagents of the craft act as if they have desires of their own, a need to
return to their dimension of origin. It would take a skilled hand to keep so
many of these in check at once. Perhaps reaching this same conclusion was what
had Bill so on edge.

\begin{dialogue}
    \item{"We're lucky, there doesn't seem to be anything other than the cauldrons that are active." Bill said, his voice a whisper.}
    \item {He beckoned me to follow him towards the large door at the far end of the room. "You said it's some kind of attuned orb, right?. This makes so much more sense now... for us to be asked, I'd assumed it was just some eccentric collector got his hands on something he shouldn't have."}
    \item {As we reached the door I told Bill, "I'm sorry I didn't tell you what I knew, I didn't think you'd let us take the job, and I'd already told Edith we'd do it."}
    \item {His rigid stance softened, "I get it.", he paused, "I don't know what we might be getting into past this door, so, if we can't speak or get separated get the orb and get out, that's all that matters. We can regroup at home."}
\end{dialogue}

He stood to the side, to let me open the door and go first. I knew this wasn't
him being a coward, quite the opposite. He knew I was going to be much more
likely to feel out where the orb was than him. I knew the feeling of magic.
Taking a deep breath, as if about to dive into a cold lake, I pushed the door
open and stepped through. What followed was a dark spiral staircase, leading
upwards. Without the ambient light and heat of the laboratory, this stairwell
felt... unwelcoming? As stupid as it sounds the lab felt homely. Proceeding
upwards we came to another innocuous door, moving through this we found ourselves
in a large entrance hall. Furnished with dark hardwoods, opulent tapestries hung
on the walls, and thick rugs covered the floor. Moving onto one of the rugs to
soften our footsteps several thoughts occurred to me. The sconces on the walls
had all gone out, usually servants would have kept these lit even during the
night. Perhaps it was because of the darkness of the room which meant it took me
a second to realise, there was quite a thick layer of dust over everything.
There were trails that made it obvious that this place was not abandoned, but it
was uncared for. This was such a strange contrast compared to the laboratory
which seemed as if it was in constant use. 

The strangeness of this place made me feel uneasy, and so I held my tongue. I
wanted to share my observations with Bill, and to hear what he'd noticed. Taking
a moment to process all these oddities I focused and reached out, feeling for
anything that might be the orb. I felt something, but it was different. Coming
from above, I felt what I could only describe as a drain. Like a plug had been
pulled. Following one of the trails of disrupted dust I could see a grand
staircase, and just beyond it a faint light, with the unmistakable flicker of
candles. Without a doubt this was the most intriguing and terrifying job I'd
ever been a part of. This felt like a tale someone would use to scare children
into behaving and not wandering off.

As I moved towards the staircase, I felt a strong hand grab my wrist.

\begin{dialogue}
    \item{"Ellen", I noted the use of my name and could sense the seriousness in his voice, "This is really off."}
\end{dialogue}

I couldn't bring myself to say something to him, even looking at him felt
outside my capabilities. All I could do was move towards the staircase. I was
caught in this drain. All concerns and fears were being sucked away too. I did
briefly wonder why Bill was still voicing concerns. Perhaps he was stronger than
me, had more mental fortitude. As I moved forward and slipped out of his grasp,
he still followed me.

Before I knew it, I was at the top of the stairs. I could see where the light
was coming from now. Before me was a large bedroom, with a four-poster bed,
and... a man. His form confused me at first. The surrounding air was distorted, it
was blurring the edges of his form. I knew at once what was happening. This
person was in the middle of planeweaving. Not a parlour trick, or simple
conjuring, this was a ritual with deep complexities. Never in my life had I
witnessed anything so beautiful and terrifying. Still I moved forward, pulled
into the room. I could see the orb now, it was being used as a focus for the
spell. I couldn't hear Bill following me now, I didn't turn around to check. 

With every breath I took in this room I could feel parts of me being stripped
away. My body felt lighter and weaker, but I kept moving towards him. This
conductor of magic, his focus was so intense that I knew he wouldn't notice me.
Closer now I saw there was something on the bed. Not a thing, a person. Laying
there so perfectly still. 

Entranced as I was, my instincts wouldn't have let me interrupt the ritual.
There was so much energy, such concentrated magic that letting it loose would be
catastrophic. Bill must have been watching such a strange sight. I stood beside
the very thing we'd been sent to steal, and the person from whom to take it, and
I just watched. 

Being surrounded by such a magic, my soul itself felt like it was being spoken
to directly. Waves of energies from dimensions beyond my comprehension hit me,
passed through me, and left me feeling like I'd just woken up for the first time
in my life. 

So lost in this strangeness I didn't even notice what happened next. I just
ceased to exist, I was a part of the infinite planes of existence.

